---
title: "Analyzing indian rice farms"
author: "Jonas Kernebeck, Alexander Nachname, Felix Lehner"
date: "07/16/2021"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\newpage

```{r imported libraries}

```

```{r constants}
# some code here
```


```{r data }
# some code here 
```

# 1 Introduction

The present data set includes production data for 171 indian rice farms. The dataframe contains the following variables:

| variable   | description                                     | expressions               |
|------------|-------------------------------------------------|---------------------------|
| id         | unique identifier for a farm                    | unique id                 |
| time       | unique identifier for a specific growing season | 1 - 6                     |
| size       | total production area in hectares               | 0.01 - 5.322              |
| status     | status of property rights                       | "owner", "share", "mixed" |
| varieties  | rice seed varietes                              | "trad", "high", "mixed"   |
| bimas      | bimas-status of the farmers                     | "no", "yes", "mixed"      |
| seed       | seed in kilogram                                | 1 - 1250 kg               |
| urea       | urea in kilogram                                | 1 - 1250 kg               |
| phosphate  | phosphate in kilogram                           | 0 - 700 kg                |
| pesticide  | pesticide cost in Rupiah                        | 0 - 62600 r               |
| pseed      | price of seed in Rupiah per kg                  | 40 - 375 r/kg             |
| purea      | price of urea in Rupiah per kg                  | 50 - 100 r/kg             |
| pphosph    | price of phosphate in Rupiah per kg             | 60 - 120 r/kg             |
| hiredlabor | hired labor in hours                            | 1 - 4536 h                |
| famlabor   | family labor in hours                           | 1 - 1526 h                |
| totlabor   | total labor (excluding harvest labor)           | 1 - 4774 h                |
| wage       | labor wage in Rupiah per hour                   | 30 - 175.35 r/h           |
| goutput    | gross output of rice in kg                      | 42 - 20960 kg             |
| noutput    | gross output minus harvesting cost              | 42 - 17610 kg             |
| price      | price of rough rice in Rupiah per kg            | 50 - 190 r/kg             |
| region     | region of the farm                              | 6 regions in Indonesia    |
\newpage


# 2 First Model

\newpage

# 3 Third Model

\newpage

# 4 Comparision

\newpage

# 5 Conclusion

\newpage